Population structure and phenotypic variation of *Sclerotinia sclerotiorum* from dry bean in the United States

### 2017-07-11 doc/manuscript/manuscript.Rmd
Add more information to results

> SEE: I was not sure what "dry bean nurseries" were because I'm use to seeing them referred to as the white mold nurseries.  I think if the introduction uses one term for it, this would be acceptable.  That or consider providing a description of the white mold nurseries in the methods section.  This now reminds me.. there is a publication by Lindsey Otto-Hanson and Steadman that gave a good description of these.  In fact, she called them multisite screening locations.  Here's the link: http://apsjournals.apsnet.org/doi/abs/10.1094/PDIS-11-10-0865

@@

## Variable Importance {-}

@@

@@

![
Barplot showing abundance of each Mycelial Compatibility Group (MCG) in descending order.
Singletons (46) have been removed for clarity.
White bars represent sample counts and grey bars represent counts of unique multilocus genotypes (MLG).
The transparency associated with the grey bars represent the evenness of the distribution of the MLGs within a given MCG as shown in the inset histograms.
](../../results/figures/publication/Figure1Z.pdf){#barplot width=50%}

@@

There were a total of 165 multilocus genotypes in the isolate collection.
The most abundant MLG represented 27 isolates (Table \@ref(tab:mlg-table)) across five regions (NE, MI, WA, CO, and ND).

@@

The associations between MCGs and MLGs are complicated (Fig. \@ref(biggraph))

@@

Community analysis of the MLGs shared between populations revealed 4 communities with a modularity of 0.17 (Fig. \@ref(community-graph)): A coastal community (CA, OR, WA, and NY), a mid-west community (CO, NE, ND, MI), and an international community (MN, France, Australia).
Analysis with 16 loci resulted in the removal of the NY node because it no longer shared a genotype with OR, but the same overall community structure was present with a modularity of 0.2 (Supplementary Figure S1Z).
The international community appears to be driven by MLG 4, which is shared between all three populations and has a $P_{sex}$ value of 2.87 x 10^-5^, in contrast to the abundant MLG 25, which has a $P_{sex}$ value of 0.0168.

## Aggressiveness {-}

The MCGs with the highest frequency were MCG 5 and MCG 44 representing 73 and 44 isolates, respectively **(Fig. 4)**. <!-- Maybe add aggressiveness information here. -->
MCGs that were unique to a region are shown in Fig. 5 while MCGs that represented at least 2 or more regions are in Fig 6. 20 MCGs were common in two or more regions and the rest were distributed uniquely in different regions.
ID, MN, NE, NY, and WI did not have any unique/isolated MCGs.
The aggressiveness of isolates ranged from 1.4 to 7.9 with a mean of 5.01.
There were six isolates that fell into upper outlier category (aggressiveness > 7.6) and one isolate in the lower outlier range (aggressiveness < 2.3) (box plot results not shown).
The mean aggressiveness of the 10 most abundant MCGs ranged from 4.6 (MCG 49) to 6.02 (MCG 44) (see Fig 4 for 10 most abundant MCGs).
ANOVA of these 10 MCGs indicated significant difference for aggressiveness levels at P <0.0001.
Mean separation of 10 most common MCGs showed MCG 44 is different from MCGs 9, 1, 45, 4, 53, and 49 while the aggressiveness levels of MCGs 2, 5, 3 failed to show any difference to MCG 44 or the other MCGs.

@@

@@

>>>>

@@

## Phenotypic variable assessment {-}

### Variable Importance {-}

@@

### Mycelial Compatibility Groups {-}

We observed 87 mycelial compatibility groups (MCG), the most abundant of which represented 73 isolates over 37 multilocus genotypes (MLG) (Fig. \@ref(barplot). 
Our assessment of aggressiveness association with MCGs showed a significant effect (P < 0.001) indicating that MCG 44 is, on average, more aggressive than other MCGs (Table \@ref(tab:mcg-aggressiveness)).

@@

We observed a total of 165 multilocus genotypes (MLG) (215 with 16 loci).
The most abundant MLG represented 27 isolates (Table \@ref(tab:mlg-table)) across five regions (NE, MI, WA, CO, and ND). 
The genotypes were relatively evenly distributed within regions with moderate to high diversity (Table \@ref(tab:iatab)). 
Of the 165 MLGs, 76 (46%) were found across all geographic regions except WI (2), ID (1), and Mexico (18) (Fig. \@ref(community-graph)).
Our graph-based approach to correlating MLGs with MCGs revealed a large and complex network (Fig. \@ref(biggraph)). 
Community analysis showed 51 communities, 23 of which consisted of a single MLG/MCG pair. 
The three communities with the most members contained 8 of the ten most abundant MCGs. 
Overall, however, there was no clear pattern to the association between MLG and MCGs.

@@

![
Barplot showing abundance of each Mycelial Compatibility Group (MCG) in descending order.
Singletons (46) have been removed for clarity, leaving 41 MCGs.
White bars represent sample counts and grey bars represent counts of unique multilocus genotypes (MLG).
The transparency associated with the grey bars represent the evenness of the distribution of the MLGs within a given MCG as shown in the inset histograms.
](../../results/figures/publication/Figure1Z.pdf){#barplot width=100%}

@@

@@

Community analysis of the MLGs shared between populations revealed 4 communities with a modularity of 0.17 (Fig. \@ref(community-graph)): A coastal community (CA, OR, WA, and NY), a mid-west community (CO, NE, ND, MI), and an international community (MN, France, Australia).
Analysis with 16 loci resulted in the removal of the NY node because it no longer shared a genotype with OR, but the same overall community structure was present with a modularity of 0.2 (Fig. \@ref(community-graph16)).
The international community appears to be driven by MLG 4, which is shared between all three populations and has a $P_{sex}$ value of 2.87 x 10^-5^, in contrast to the abundant MLG 25, which has a $P_{sex}$ value of 0.0168.

@@

```{r table2, echo = FALSE, warning = FALSE, message = FALSE, results = "asis"}
suppressPackageStartupMessages(library("tidyverse"))
read_csv("../../results/tables/MCG-aggressiveness.csv", 
         na = "NA", # spaces read in as spaces
         col_types = cols(.default = col_character())) %>%
  dplyr::rename(MCG = trt) %>%
  dplyr::rename(Group = M) %>%
  dplyr::rename(`Mean Aggressiveness` = means) %>%
  huxtable::as_huxtable(add_colnames = TRUE) %>%
  huxtable::set_number_format(row = huxtable::everywhere, col = 1, 0) %>% 
  huxtable::set_col_width(c(0.05, 0.3, 0.08)) %>%
  huxtable::set_align(row = huxtable::everywhere, col = 2, "right") %>%
  huxtable::print_md(max_width = 31)
```

Table: (\#tab:mcg-aggressiveness) Mean aggressiveness ratings for the ten most abundant mycelial compatibility groups (MCG) with groupings according to 95% family-wise confidence interval.

\newpage

@@

![
Network of populations (nodes/circles) and their shared multilocus haplotypes (MLG) (edges/lines) genotyped over 16 loci. 
Each node is labeled with **name (number of MLGs shared/number of MLGs total).** 
The shade and area of the nodes are proportional to the number of unique MLGs within the node and the inner nodes are proportional to the number of private MLGs to the region. 
Each edge represents a single MLG where its thickness represents the number of populations that share the MLG and the shade represents the value of $P_{sex}$, or the probability of encountering that MLG from two independent meiotic events.
](../../results/figures/publication/FigureS1Z.pdf){#community-graph16 width=100%}

\newpage
